---
title: "DQR Project2"
author: "Cheng Chen"
date: "3/4/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: File Description
File name: Application dataset
Source (url, client…): 
Number of records: 100,000
Number of fields: 9 fields, 1 index, 2 date, 6 categorial
Time frame: 01/01/2015 - 12/31/2015

```{r cars}
data = read.csv("Application.csv")
library(lubridate)
library(ggplot2)
library(dplyr)
```

## Part 2: List of info for each field

Description (continuous, categorical with metric, categorical no metric)
% populated
# unique values
Min, max, mean, median, mode, standard deviation for continuous
Picture (either a distribution, histogram, table…)


### Field 1: record
Description: Index of record
Number of unique values: 100,000, start from 1, interval is 1, no missing value

### Field 2: date
Description: date of each application record, categorical with metric
Percent of Populated: 100%, no missing values
Number of unique values: 365, from 01/01/2015 to 12/31/2015

```{r Field 2, echo=FALSE}
# transfer to standard date style
data$date = ymd(data$date)
# Number of unique values
length(unique(data$date))
# plot the histogram by day
ggplot(data, aes(x = date)) + 
  geom_histogram(bins = 365, color = "steelblue") + 
  ggtitle("Distribution of Application by Day") + 
  theme_bw() + 
  xlab("Application Date")
# plot the histogram by week
date0 = data.frame(as.factor(week(data$date)))
names(date0) = c("date0")
ggplot(date0, aes(x = date0)) +
  geom_bar(fill = "steelblue") +
  ggtitle("Distribution of Application by Week") + 
  theme_bw() +
  xlab("Application Week")
# plot the histogram by day of week (on average) ## 但是这里不是on average的结果
date1 = data.frame(as.factor(wday(data$date, label = T, abbr = F)))
names(date1) = c("date1")
ggplot(date1, aes(x = date1)) +
  geom_bar(fill = "steelblue") +
  ggtitle("Distribution of Application by Day of Week") + 
  theme_bw() + 
  xlab("Day of Week")
# plot the histogram by month
date2 = data.frame(as.factor(month(data$date, label = T, abbr = T)))
names(date2) = c("date2")
ggplot(date2, aes(x = date2)) +
  geom_bar(fill = "steelblue") +
  ggtitle("Distribution of Application by Month") + 
  theme_bw() + 
  xlab("Application Month")
```

### Field 3: ssn
Description: ssn of each application record, categorical with no metric
Percent of Populated: 100%, no missing values
Number of unique values: 96535

```{r Field 3, echo=FALSE}
# Number of unique values
length(unique(data$ssn))
# Find the most 20 frequently used ssn
ssn = data %>%
  group_by(ssn) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
ssn1 = ssn[1:20, ]
ggplot(ssn1, aes(x = reorder(ssn, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 SSNs") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("SSN")

max(data$ssn) ## 9 9999 3079
min(data$ssn) ## 2503
ssn_f1 = data.frame(data$ssn)
ssn_f = ssn_f1[ssn_f1 < 1e8]
length(ssn_f)
ssn_f = data.frame(ssn_f)
names(ssn_f) = "ssn"
View(ssn_f)

ssn_f1 = ssn_f %>%
  group_by(ssn) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
ssn_ff = ssn_f1[1:20, ]
ggplot(ssn_ff, aes(x = reorder(ssn, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 Frivolous SSNs") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("SSN")
```


### Field 4: firstname
Description: firstname of each application record, categorical with no metric
Percent of Populated: 100%, no missing values
Number of unique values: 16576

```{r Field 4, echo=FALSE}
# Number of unique values
length(unique(data$firstname))
# Find the most 20 frequently used firstname
firstname = data %>%
  group_by(firstname) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
firstname1 = firstname[1:20, ]
ggplot(firstname1, aes(x = reorder(firstname, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 Firstnames") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("Firstname")
```

### Field 5: lastname
Description: lastname of each application record, categorical with no metric
Percent of Populated: 100%, no missing values
Number of unique values: 36312

```{r Field 5, echo=FALSE}
# Number of unique values
length(unique(data$lastname))
# Find the most 20 frequently used lastname
lastname = data %>%
  group_by(lastname) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
lastname1 = lastname[1:20, ]
ggplot(lastname1, aes(x = reorder(lastname, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 Lastnames") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) + 
  xlab("Lastname")
```

## Fullname 

Description: Fullname of each application record, categorical with no metric
Percent of Populated: 100%, no missing values
Number of unique values: 93726
```{r Field 5.5, echo=FALSE}
## Plot the most 20 frequent name (first + last)
name = data.frame(paste(data$firstname, data$lastname, sep = " "))
names(name) = c("Name")
# Number of unique values
length(unique(name$Name))
Fullname = name %>%
  group_by(Name) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
Fullname1 = Fullname[1:20, ]
ggplot(Fullname1, aes(x = reorder(Name, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 Fullnames") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("Full Name")
```

### Field 6: address
Description: address of each application record, categorical with no metric
Percent of Populated: 100%, no missing values
Number of unique values: 97563
Frivolous item: 2602 AJTJ AVE
```{r Field 6, echo=FALSE}
# Number of unique values
length(unique(data$address))
# Find the most 20 frequently used address
address = data %>%
  group_by(address) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
address1 = address[1:20, ]
ggplot(address1, aes(x = reorder(address, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 Addresses") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("Address")
```


### Field 7: ZIP5
Description: zip5 of each application record, categorical with no metric
Percent of Populated: 100%, no missing values
Number of unique values: 16547
```{r Field 7, echo=FALSE}
# Number of unique values
length(unique(data$zip5))
# Find the most 20 frequently used zip5
zip5 = data %>%
  group_by(zip5) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
zip51 = zip5[1:20, ]
ggplot(zip51, aes(x = reorder(zip5, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 ZIP5") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("ZIP")

max(data$zip5) ## 99999
min(data$zip5) ## 2
zip_f1 = data.frame(data$zip5)
zip_f = zip_f1[zip_f1 < 1e4]
length(zip_f)
zip_f = data.frame(zip_f)
names(zip_f) = "zip"

zip_ff1 = zip_f %>%
  group_by(zip) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
zip_ff = zip_ff1[1:20, ]
ggplot(zip_ff, aes(x = reorder(zip, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 Frivolous Zips") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("Frivolous Zips")
```

### Field 8: dob
Description: birth date of each applicant, categorical with metric
Percent of Populated: 100%, no missing values
Number of unique values: 36816, from 01/01/1900 to 10/31/2016

```{r Field 8, echo=FALSE}
# transfer to standard date style
data$dob = ymd(data$dob)
# summary of dob
length(unique(data$dob))
max(data$dob)
min(data$dob)
# Find the most 20 frequently used DOB
dob = data %>%
  group_by(dob) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
dob1 = dob[1:20, ]
ggplot(dob1, aes(x = reorder(dob, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 Applicant DOB") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) + 
  xlab("DOB")
# plot the histogram by week
dob2 = data.frame(as.factor(week(data$dob)))
names(dob2) = c("dob2")
ggplot(dob2, aes(x = dob2)) +
  geom_bar(fill = "steelblue") +
  ggtitle("Distribution of Applicant DOB by Week") + 
  theme_bw() + 
  xlab("Birth Week")
# plot the histogram by day of week (on average) ## 但是这里不是on average的结果
dob3 = data.frame(as.factor(wday(data$dob, label = T, abbr = F)))
names(dob3) = c("dob3")
ggplot(dob3, aes(x = dob3)) +
  geom_bar(fill = "steelblue") +
  ggtitle("Distribution of Applicant DOB by Day of Week") + 
  theme_bw() +
  xlab("Birth Day of Week")
# plot the histogram by month
dob4 = data.frame(as.factor(month(data$dob, label = T, abbr = T)))
names(dob4) = c("dob4")
ggplot(dob4, aes(x = dob4)) +
  geom_bar(fill = "steelblue") +
  ggtitle("Distribution of Applicant DOB by Month") + 
  theme_bw() + 
  xlab("Birth Month")
```


### Field 9: homephone
Description: homephone of each application record, categorical with no metric
Percent of Populated: 100%, no missing values
Number of unique values: 22181
```{r Field 9, echo=FALSE}
# Number of unique values
length(unique(data$homephone))
# Find the most 20 frequently used homephone
homephone = data %>%
  group_by(homephone) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
homephone1 = homephone[1:20, ]
ggplot(homephone1, aes(x = reorder(homephone, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 homephone") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("Homephone")


max(data$homephone)
min(data$homephone)

max(data$homephone) ## 99999
min(data$homephone) ## 2
homephone_f1 = data.frame(data$homephone)
homephone_f = homephone_f1[homephone_f1 < 1e9]
length(homephone_f)
homephone_f = data.frame(homephone_f)
names(homephone_f) = "homephone"

homephone_ff1 = homephone_f %>%
  group_by(homephone) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency))
homephone_ff = homephone_ff1[1:20, ]
ggplot(homephone_ff, aes(x = reorder(homephone, -frequency), y = frequency)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
  coord_trans(y = "sqrt") + 
  ggtitle("Most Frequent 20 Frivolous Homephones") + 
  geom_text(aes(label = frequency, y = frequency), size = 3, vjust = -0.5) +
  xlab("Frivolous Homephone")
```

