---
title: "R Variables Creation"
output: github_document
---


#### 0. library and read dataset
```{r library}
## Read dataset
data = read.csv("Application.csv")
## Load library
library(lubridate)
library(ggplot2)
library(dplyr)
```

#### 1. Add three new variable: full name, person, yearday
```{r NewVariables}
#  Create three new variables
## fullname = firstname + " " + lastname
data$fullname = as.factor(paste(data$firstname, data$lastname, sep = " ")) 
## target person = fullname + " " + dob
data$person = as.factor(paste(data$fullname, data$dob, sep = " "))
## yearday = yday
data$yearday = yday(ymd(data$date))
```

#### 2. Dataset attributes
```{r DataSetAttributes}
## Number of variables in current dataset
n = ncol(data)
## set time interval, 3, 7, 14, 21 days
## calculate number of combinations observed in the past "DateInt" days
DateInt = c(3, 7, 14, 21)
```

We focus on five variables and their associations:
Column3: ssn; Column6: address; Column7: zip5; Column9: homephone; Column11: person
Yearday is used for calculation: 
Column12: yearday  

#### 3.1  Full loop for calculating times that each variable occurs in the past DataInt days
the occurrence of record itself is not included
exclude Column7(zip5) in this computation
```{r PastDateIntNumber}
data_four_sin = data[,c(3,6,9,11,12)] 
n_d_s = ncol(data_four_sin) - 1
for(s in 1:4){
  for(i in 1:(n_d_s)){
    data_ssn_larger_one = subset(assign(paste("data", "_", names(data_four_sin)[i], sep = ""), data_four_sin %>%
    group_by(data_four_sin[,i]) %>% 
    summarise(num_each = n()) %>%
    arrange(-num_each)), num_each>1)[,1]
    n_exist = ncol(data_four_sin)
    data_four_sin[,n_exist+1] = 0
    colnames(data_four_sin) = c(colnames(data_four_sin[1:n_exist]), paste("NumLast", names(data_four_sin)[i], "Int", DateInt[s], sep = ""))
    for(j in 1:length(unlist(data_ssn_larger_one))){
        data_index_each = which(data_four_sin[,i] == unlist(data_ssn_larger_one[,1])[j])
        data_index_each_date = data_four_sin$yearday[data_index_each]
        n_each = length(data_index_each)
        for(k in 1:n_each){
          data_four_sin[,n_exist+1][data_index_each[k]] = sum(data_index_each_date>=(data_index_each_date[k]-DateInt[s]+1) &
                                                      data_index_each_date<=data_index_each_date[k]) - 1
}}}}
data = data.frame(data,data_four_sin[,(n_d_s+2):(5*n_d_s+1)])
```

#### 3.2  Full loop for calculating times that association of two variables occurs in the past DataInt days
the occurrence of record itself is not included
ssn, address, zip5, homephone, person, 10 combinations in total
```{r PastDateIntNumberCombination}
data_side = data[,c(3,6,3,7,3,9,3,11, 6,7,6,9,6,11, 7,9,7,11, 9,11, 12)] 
## 5C2 = 10  ## length = 21
ncol_ds = ncol(data_side)
for(s in 1:4){
  for(i in 1:((ncol_ds-1)/2)){
    data_ssn_larger_one = subset(assign(paste("data", "_", names(data_side)[2*i-1], names(data_side[2*i]), "Int", DateInt[s], sep = ""), data_side %>%
    group_by(data_side[,2*i-1], data_side[,2*i]) %>% 
    summarise(num_each = n()) %>%
    arrange(-num_each)), num_each>1)[,1:2]
    n_exist = ncol(data_side)
    data_side[,n_exist+1] = 0
    colnames(data_side) = c(colnames(data_side[1:n_exist]), paste("NumLast", names(data_side)[2*i-1], names(data_side[2*i]), "Int", DateInt[s], sep = ""))
    for(j in 1:length(unlist(data_ssn_larger_one))/2){
        data_index_each = which(data_side[,2*i-1] == unlist(data_ssn_larger_one[,1])[j] & data_side[,2*i] == unlist(data_ssn_larger_one[,2])[j])
        data_index_each_date = data_side$yearday[data_index_each]
        n_each = length(data_index_each)
        n_each = ifelse(n_each<2, 2, n_each) 
        for(k in 1:n_each){
          data_side[,n_exist+1][data_index_each[k]] = sum(data_index_each_date>=(data_index_each_date[k]-DateInt[s]+1) &
                                                      data_index_each_date<=data_index_each_date[k]) - 1
}}}}
data = data.frame(data,data_side[(ncol_ds+1):(ncol_ds+4*(ncol_ds-1)/2)]) 
```

#### 4.1  # of days since last saw this ssn, address, homephone, person 
exclude Column7(zip5) in this computation
those records without "another occurrence" are assigned with 999.
```{r LastOccurDay}
data_four_sin = data[,c(3,6,9,11,12)]
n_data_four_sin = ncol(data_four_sin)
for(i in 1:(ncol(data_four_sin)-1)){
  data_ssn_larger_one = subset(assign(paste("data", "_", names(data_four_sin)[i], sep = ""), data_four_sin %>%
  group_by(data_four_sin[,i]) %>% ## ??????????????????????????????
  summarise(num_each = n()) %>%
  arrange(-num_each)), num_each>1)[,1]
  n_exist = ncol(data_four_sin)
  data_four_sin[, n_exist+1] = 999 ## i will prefer 365*2 = 730 here
  colnames(data_four_sin) = c(colnames(data_four_sin[1:n_exist]), paste("NumLast", names(data_four_sin)[i], sep = ""))
  for(j in 1:length(unlist(data_ssn_larger_one))){
      data_index_each = which(data_four_sin[,i] == unlist(data_ssn_larger_one[,1])[j])
      data_index_each_date = data_four_sin$yearday[data_index_each]
      n_each = length(data_index_each)
      for(k in 2:n_each){
        data_four_sin[,n_exist+1][data_index_each[k]] = data_index_each_date[k]-data_index_each_date[k-1]
}}}
data = data.frame(data,data_four_sin[,(n_data_four_sin+1):(n_data_four_sin+4)])
```

#### 4.2  # of days since last saw this combinations of two variables
ssn, address, zip5, homephone, person, 10 combinations in total
those records without "another occurrence" are assigned with 999.
```{r LastOccurDayCombination}
data_four = data[,c(3,6,3,7,3,9,3,11, 6,7,6,9,6,11, 7,9,7,11, 9,11, 12)] 
## 5C2 = 10  ## length = 21
ncol_ds = ncol(data_four)
for(i in 1:((ncol_ds-1)/2)){
  data_ssn_larger_one = subset(assign(paste("data", "_", names(data_four)[2*i-1], names(data_four[2*i]), sep = ""), data_four %>%
  group_by(data_four[,2*i-1], data_four[,2*i]) %>% 
  summarise(num_each = n()) %>%
  arrange(-num_each)), num_each>1)[,1:2]
  n_exist = ncol(data_four)
  data_four[, n_exist+1] = 999 
  colnames(data_four) = c(colnames(data_four[1:n_exist]), paste("NumLast", names(data_four)[2*i-1], names(data_four[2*i]), sep = ""))
  for(j in 1:length(unlist(data_ssn_larger_one))/2){
      data_index_each = which(data_four[,2*i-1] == unlist(data_ssn_larger_one[,1])[j] & data_four[,2*i] == unlist(data_ssn_larger_one[,2])[j])
      data_index_each_date = data_four$yearday[data_index_each]
      n_each = length(data_index_each)
      n_each = ifelse(n_each<2, 2, n_each) 
      for(k in 2:n_each){
        data_four[,n_exist+1][data_index_each[k]] = data_index_each_date[k]-data_index_each_date[k-1]
}}}
data = data.frame(data,data_four[(ncol_ds+1):(ncol_ds+(ncol_ds-1)/2)]) 
```


#### 5. Modify the Frivolous Data
```{r FrivolousData}
## discrete the dataset
data1 <- data[,1:12]
data2 <- data[,13:68]
data3 <- data[,69:82]
prep_data <- cbind(data1, data2, data3)

## Calculate the means without the frivolous
holder1 <- prep_data %>% 
  filter(dob != c(19070626), address != c("2602 AJIT AVE"), ssn != 737610282, homephone != 9105580920) 
m = NULL
for(i in 13:82){m <- c(m, mean(holder1[,i]))}

## Replace the frivolous value
holder2 <- prep_data %>% 
  filter(dob == c(19070626) | address == c("2602 AJIT AVE") | ssn == 737610282 | homephone == 9105580920) 
n <- a2[1:nrow(holder2),13:82]
for(i in 1:70){n[,i] <- m[i]}
holder2[13:82] <- n
final_data <- rbind(holder1, holder2)
## reorder according to record..
final_data <- arrange(final_data, record..)
save(final_data, file = "final_data.rda")
```

#### 6. Remove variables that correlation == 1
```{r BeforePCA}
load("full_data.rda")
final1 = final_data[,c(13:82)]
correlation1 = as.data.frame(cor(final1))
length(which(correlation1 == 1)) 

final = final_data[,c(13:27, 29:49, 51:58, 60:61, 63:64, 66, 69:78, 80:82)] 
correlation = as.data.frame(cor(final))
length(which(correlation == 1)) == ncol(final) ## TRUE
```


